---
title: "Checking control flow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking control flow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

You can use `check_for()`, `check_while()` and `check_if_else()` chains to check whether a student has properly coded control constructs.

## Example 1: For loop

Consider the following solution:

```{r}
# Print out the integers 1 to 10
for (i in 1:10) {
  print(i)
}
```

A suitable SCT for this exercise can be the following:

```{r}
forloop <- ex() %>% check_for(not_found_msg = "Make sure to code a for loop in the first place!")
forloop %>% check_cond() %>% check_code("in\\s*1:10", missing_msg = "You can use `i in 1:10` to define your for loop.")
forloop %>% check_body() %>% check_function("print") %>% check_arg("x") %>% check_equal(eval = FALSE)
```

- `check_for()` checks whether students have specified a `for` loop in their code. If this is the case, it produces a substate, referring to the condition of the for loop and its body.
- Next, you can 'zoom in' on the condition and body of the `for` loop, using `check_cond()` and `check_body()` respectively.  After this zooming in, you can continue the chain of SCT functions to check these 'parts' of the code. The `check_code()` call, for example, only looks to match the regular expression in the iteration part of the `for` loop.

## Example 2: Check if else

Consider the following solution:
	
```{r}
# Predefined value of x
x <- TRUE

# Code the if else construct
if (x) {
  print("x is TRUE!")
} else {
  print("x is not TRUE!")
}
```

The following SCT checks its correctness:

```{r}
ex() %>% check_object("x")
ifelse <- ex() %>% check_if_else()
ifelse %>% check_cond() %>% check_code("x")
ifelse %>% check_if() %>% check_function("print") %>% check_arg("x") %>% check_equal()
ifelse %>% check_else() %>% check_function("print") %>% check_arg("x") %>% check_equal()
```

- `check_if_else()` parses the student code and checks if there is an if-else statement in there.
- `check_cond`, `check_if`, and `check_else()` all 'zoom in' on a particular part of the if-else statement: the condition, the body of the `if` statement and the body of the `else` statement. After this zooming in, you can continue the chain of SCT functions to check these 'parts' of the code. The first `check_function("print")` chain for example, only looks for the function call inside the `if` body.

### Example 3: Check if, else if, else

It is also possible to use `else if` in R. How to test this? Well, behind the scenes, R parses the following structure:

```{r}
if (condition1) {
  expression
} else if (condition2) {
  expression
} else {
  expression
}
```

as if it follows this structure:

```{r}
if (condition1) {
  expression
} else {
  if (condition2) {

  } else {

  }
}
```

If you want to test such a piece of code, you therefore need the following construct:

```{r}
ex() %>% check_if_else() %>% {
  check_cond(.) %>% ...
  check_if(.) %>% ...
  check_else(.) %>% check_if_else() %>% {
    check_cond(.) %>% ...
    check_if(.) %>% ...
    check_else(.) %>% ...
  }
}
```

## Caution

Not all of the SCT chains that you specify after a 'zooming in' step might work as you'd expect. Remind that functions like `test_object()` and `test_function()` often depend on both the student and solution environment. It is possible that during execution of control flow these values change. The `testwhat` functions will always compare the 'end environment' of student and solution, it is not possible to do matching on intermediate values of objects that are changed further in the script.

