---
title: "Checking function definitions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking function definitions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Some exercises require a student to hand-craft a function to make their code DRY. With `check_fun_def()` and a range of related helper `check_` functions you can check the correctness of these user-defined functions.

As an example, have a look at the following solution:

```{r}
# Define my_fun
my_fun <- function(a, b) {
  abs(a) + abs(b)
}
```

The following SCT checks the function definition and whether the function was properly called:

```{r}
ex() %>% check_fun_def("my_fun") %>% {
  check_arguments(.)
  check_call(., 1, 2) %>% check_result() %>% check_equal()
  check_call(., -1, 2) %>% check_result() %>% check_equal()
  check_call(., 1, -2) %>% check_result() %>% check_equal()
  check_call(., -1, -2) %>% check_result() %>% check_equal()
  check_body(.) %>% {
    check_function(., "abs", index = 1)
    check_function(., "abs", index = 2)
  }
}
```

- `check_fun_def` checks whether `my_fun` was defined by the student. If this is the case, it will produce a substate that contains a reference to the function arguments and its body.
- `check_arguments()` checks whether the student specified the same number of arguments as the function definition in the solution did.
- `check_call()` will create an R expression that calls the function `my_fun()` with the arguments specified in it. As an example, `check_call(., 1, 2)` will generate the expression `my_fun(1, 2)`. `check_result()` will then execute this expression (in both student and solution workspace), and focus on the result of the expression. Finally, `check_equal()` verifies whether these call results correspond.
- `check_body()` zooms in on the body of the function definition (as if the body of the function was the only code in the student submission). If, for example, the student used `abs()` outside the function definition, this would be ignored. Notice here that no arguments are specified. If we did this, `check_function()` would compare the actual values of the arguments. This is information that is not available, as `a` and `b` are not variables that are available in the workspace, only in the scope of the function.

**Note:** in addition to `check_result()`, you can also pipe the result of `check_call()` into `check_output()` and `check_error()`, and follow up with `check_equal()`. These functions will respectively check the output and error the function call generated when executed in the student/solution environment.
