---
title: "Syntax"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Syntax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`testwhat` uses the pipe operator (`%>%`) from the [`magrittr`](https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html) package to 'chain together' SCT functions. Every chain starts with the `ex()` function call, which holds the exercise state. This exercise state contains all the information that is required to check if an exercise is correct, which are:

+ the student submission and the solution as text, and their corresponding parse trees.
+ a reference to the student workspace and the solution workspace.
+ the output and errors that were generated when executing the student code.

As SCT functions are chained together with `%>%`, the exercise state is copied and adapted into so-called child states to zoom in on particular parts of the code.

## Example

Consider the following snippet of markdown that represents part of an exercise:

    `@solution`
    ```{r}
    x <- 4
    if (x > 0) {
      print("x is positive")
    }
    ```
    
    `@sct`
    ```{r}
    if_state <- ex() %>% check_if_else() %>% {
        check_cond(.) %>% check_code(c("x\\s+>\\s+0", "0\\s+<\\s+x")) # chain A
        check_if(.) %>% check_function("print") %>% check_arg("x") %>% check_equal() # chain B
    }
    ```

- `check_if_else()` will check whether an `if` statement was coded, and will afterwards 'zoom in' on the if statement only.
- Chain A: `check_cond()` will consequently zoom in on the condition part of the `if` statement. Hence, `check_code()` will only look inside the condition of the `if` statement of the student's submission.
- Chain B: Similarly, `check_if()` starts from the `if` statement, and zooms in on the body of the `if` statement, after which `check_function()` will only look for the `print` call inside the body, ignoring anything outside of it.

To further explain this example, assume the following student submission:

```{r}
x <- 4
if (x < 0) {
  print("x is negative")
}
```

In chain A, this is what happens:

- `check_if_else()` considers the entire submission (as contained in `ex()`), and produces a 'child state' that contains the `if` statements in student and solution:

  ```{r}
  # solution
  if (x > 0) {
    print("x is positive")
  }
  
  # student
  if (x < 0) {
    print("x is negative")
  }
  ```

- `check_cond()` considers the state above produced by `check_if_else()`, and produces a 'child state' with only the condition parts of the `if` statements:

  ```{r}
  # solution
  x > 0
  
  # student
  x < 0
  ```

- `check_code()` considers the state above produced by `check_cond()`, and tries to match the regexes to `x < 0` in the student. None of the regexes match this fragment of the student code, so the test fails.

Assume now that the student corrects the mistake, and submits the following, which is still not correct:

```{r}
x <- 4
if (x > 0) {
  print("x is negative")
}
```

Chain A will go through the same steps and will pass this time as `x > 0` is matches one of the regexes. In Chain B, this is what happens:

- `check_if()` considers the state produced by `check_if_else()`, and produces a 'child state' with only the body parts of the `if` statements:

  ```{r}
  # solution
  print("x is positive")
  
  # student
  print("x is negative")
  ```

- `check_function()` considers the state above produced by `check_if()`, and tries to find the function `print()`. Next, it produces a state with references to the different arguments that were specified and their values:

  ```
  # solution
  { "x": "x is positive" }

  # student
  { "x": "x is negative" }
  ```
  
- `check_arg()` checks if the argument `x` is specified, and produces a child state that zooms in on the actual value of `x`:

  ```r
  # solution
  "x is positive"

  # student
  "x is negative"
  ```
  
- Finally, `check_equal()` compares the equality of the two 'focused' arguments. They are not equal, so the check fails.

