---
title: "Test to check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Test to check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

If you are looking at the SCTs of old DataCamp courses, you'll notice they use `test_x()` functions instead of `check_x()` functions, and there is no usage of `ex()`. The `test_x()` way of doing things has now been phased out in terms of the more verbose, but more transparent and composable `check_x()` functions that start with `ex()` and are chained together with the `%>%` operator.

Whenever you come across an SCT that uses `test_x()` functions, you'll make everybody's life easier by converting it to a `check_x()`-based SCT. Below are the most common cases you will encounter, together with instructions on how to translate from one to the other.

Something you came across that you didn't find in this list? Just create an issue on [`testwhat`'s GitHub repo](https://github.com/datacamp/testwhat). Content Engineering will explain how to translate the SCT and update this article.

### `test_student_typed()`

```{r}
# Solution
TRUE & FALSE

# old SCT (fixed = TRUE by default)
test_student_typed(c("TRUE & FALSE", "FALSE & TRUE"))

# new SCT (fixed = FALSE by default)
ex() %>% check_code(c("TRUE & FALSE", "FALSE & TRUE"), fixed = TRUE)
```

```{r}
# Solution
"Hello, world!"

# equivalent 'old' SCT (fixed = TRUE by default)
test_student_typed("[H|h]ello,*\\s*[W|w]orld\\!*", fixed = FALSE)

# SCT, robust to small typos
ex() %>% check_code("[H|h]ello,*\\s*[W|w]orld\\!*")
```

### `test_object()`

```{r}
# Solution
x <- 5

# old SCT
test_object("x")

# new SCT
ex() %>% check_object("x") %>% check_equal()
```

```{r}
# Solution, value of x does not matter
x <- 5

# old SCT
test_object("x", eval = FALSE)

# new SCT
ex() %>% check_object("x")
```

### `test_function()`

```{r}
# Solution
mean(1:3)

# old SCT
test_function("mean", args = c("x"))

# new SCT
ex() %>% check_function("mean") %>% check_arg("x") %>% check_equal()
```

```{r}
# Solution, value of na.rm does not matter
mean(c(1, 2, 3), na.rm = TRUE)

# Old SCT
test_function("mean", args = c("x", "na.rm"), eval = c(TRUE, NA))

# SCT
ex() %>% check_function("mean") %>% {
  check_arg(., "x") %>% check_equal()
  check_arg(., "na.rm")
}
```

### `test_error()`

```{r}
# old SCT
test_error()

# new SCT
ex() %>% check_error()
```

### `test_function_result()`

```{r}
# Solution
mean(1:3)

# old SCT
test_function_result("mean")

# new SCT
ex() %>% check_function("mean") %>% check_result() %>% check_equal()
```

### `test_correct()`

Still use `test_correct()` for now.

```{r}
# Solution (how x is calculated does not matter)
x <- mean(1:3)

# old_SCT
test_correct(test_object("x"), test_function("mean", args = c("x")))

# new SCT
test_correct(
  ex() %>% check_object("x") %>% check_equal(),
  ex() %>% check_fun("mean") %>% check_arg("x") %>% check_equal()
)
```

